\documentclass{article}
\usepackage{iclr2026_conference,times}

\input{math_commands.tex}

\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{url}
\usepackage{xcolor}
\usetikzlibrary{arrows.meta,backgrounds,calc,fit,matrix,positioning,shapes.geometric}

\definecolor{llblue}{HTML}{4E79A7}
\definecolor{llorange}{HTML}{F28E2B}
\definecolor{llgreen}{HTML}{59A14F}
\definecolor{llred}{HTML}{E15759}
\definecolor{llpurple}{HTML}{B07AA1}
\definecolor{llgray}{HTML}{6B7280}
\definecolor{lllight}{HTML}{F4F6F8}

\newcommand{\code}[1]{\texttt{#1}}

\title{VeriAnim: Verifiable LLM 3D Animation Generation with Blender Programs}

\author{Anonymous Authors\\
Anonymous Institution\\
\texttt{anonymous@example.com}
}

%\iclrfinalcopy
\begin{document}

\maketitle

\begin{abstract}
LLM 3D animation generation needs static and temporal contracts. Blender programs give language models an inspectable medium for constructing and animating 3D scenes. Animation adds obligations that static generation does not expose: object identities must persist, events must own frame windows, contacts must hold across interpolation, cameras must reveal the motion, and unsupported dynamics must fail with clear scope. We present VeriAnim, a static SceneSpec IR, an animation-extension layer, and VeriAnim-AnimBench for verifiable LLM-generated Blender animations. SceneSpec represents object identity, parts, materials, relations, cameras, collision proxies, and evidence plans before code generation. The harness first verifies and repairs the static scene with deterministic checks, screenshots, and visual verification. The animation layer then extends that verified baseline with rigid motion, camera motion, visibility changes, and deformable events. The harness records scene graphs and media artifacts, queries visual and video verifiers, and applies bounded repair when measured evidence identifies a local violation. VeriAnim-AnimBench contributes 300 natural-language prompts across easy, medium, and hard animation tiers.
\end{abstract}

\section{Introduction}

Blender has become a programmable medium for communicating dynamic 3D ideas: demonstrations, robotics mockups, educational animations, game prototypes, and synthetic data scenes all rely on motion as much as shape. A useful LLM 3D animation interface should therefore produce an executable animation, not just a static arrangement of objects. A prompt such as ``a gripper moves a box from a conveyor to a cart while the camera follows'' requires object construction, event sequencing, contact maintenance, camera coverage, and temporal verification.

Blender Python gives this interface a practical target language for large language 3D modeling~\citep{lu2025verianim}. Static scene generation already benefits from code: object creation, materials, lights, cameras, and layout constraints become inspectable program structure. Static scenes still need their own verifier-repair loop because animation should start from measured object identities, relations, camera coverage, and material evidence. We add a static SceneSpec IR that makes those obligations explicit before code generation. Animation exposes the next structural gap. A static scene contract says what exists, what relation evidence supports it, and what repairs preserve it. An animation contract must say which object moves, who controls each frame window, which relation persists, which evidence verifies the motion, and which runtime executes the behavior.

Direct keyframe generation collapses these obligations into unstructured code. The generated script may satisfy the final frame while interpolating through a support surface, move a payload while leaving the gripper behind, or render a video in which the important contact point is hidden. These outcomes do not mean Blender code is the wrong medium. They show that animation needs a contract layer between language planning, code generation, execution, verification, and repair.

We introduce a two-level IR for verifiable Blender generation. \code{SceneSpec} captures the static scene as a typed contract over objects, parts, materials, relations, cameras, proxies, and evidence. The \emph{Animation Extension Contract} then keeps that scene as a frozen baseline and adds animation families, event windows, target capability profiles, simulation-cache metadata, and verifier probes. Rigid-body motion uses deterministic geometry probes. Camera and visibility events use coverage and media probes. Deformable events use deformation statistics plus video evidence. Capability profiles record runtime support and unsupported scope.

The design follows a simple rule: every generated scene or animation feature must name the evidence that can validate it. A planner emits a scene-and-animation IR. A coder generates Blender Python against a stable helper library. A Blender addon executes the script and returns scene graphs, transform traces, screenshots, and preview videos. Deterministic validators check support, penetration, containment, attachment, baseline preservation, and transform ownership. Visual and video verifiers judge semantic layout, visibility, and temporal intent. Repair routines rewrite local scene placements or motion paths when measured evidence identifies a bounded violation.

This paper makes four contributions:
\begin{itemize}
    \item \textbf{Static and temporal IR contracts.} We define \code{SceneSpec} for static Blender scenes and an animation extension that organizes families, probes, capability profiles, event windows, and artifact ownership.
    \item \textbf{Verifier-facing Blender harness.} We use Blender code as a 3D medium and add staged static-to-animation generation, static and temporal repair, visual/video verification, and media diagnostics.
    \item \textbf{Rigid primitive methodology.} We instantiate the contract with geometry-aware rigid primitives for support motion, paired interaction, pick-place manipulation, articulation, and temporal composition.
    \item \textbf{VeriAnim-AnimBench.} We release a 300-prompt benchmark with easy, medium, and hard tiers for measuring scene complexity, temporal composition, verifier coverage, and repair cost.
\end{itemize}
The implementation supports static scene generation, camera and rigid animation events, texture resolution, visual/video verifier loops, deformable Blender events, capability-profile reporting, and benchmark generation.

\begin{figure}[t]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
    font=\small,
    stage/.style={draw, rounded corners=5pt, thick, align=center, minimum width=38mm, minimum height=8mm, inner sep=4pt},
    card/.style={draw, rounded corners=3pt, align=center, minimum width=35mm, minimum height=10mm, inner sep=4pt, fill=white},
    artifact/.style={draw, rounded corners=3pt, align=center, minimum width=34mm, minimum height=7mm, inner sep=3pt, fill=lllight},
    arrow/.style={-Latex, thick},
    faintarrow/.style={-Latex, thick, llgray},
    repair/.style={-Latex, thick, dashed, llred}
]
\node[stage, fill=llblue!12] (s1) at (0,0) {1. Plan};
\node[stage, fill=llorange!14] (s2) at (4.4,0) {2. Generate static scene};
\node[stage, fill=llpurple!14] (s3) at (8.8,0) {3. Add animation};
\node[stage, fill=llgreen!14] (s4) at (13.2,0) {4. Verify animation};

\node[card] (prompt) at (0,-1.25) {Natural-language\\prompt};
\node[card, fill=llblue!5] (planner) at (0,-2.65) {Planner\\SceneSpec + extension};
\node[card, fill=llorange!6] (scene) at (4.4,-1.25) {Static coder\\objects, materials, cameras};
\node[artifact] (staticart) at (4.4,-2.65) {Scene graph\\screenshots + reports};
\node[card, fill=llorange!10, minimum width=38mm] (staticfix) at (4.4,-4.05) {Static repair\\or LLM refinement};
\node[card, fill=llpurple!6] (anim) at (8.8,-1.25) {Animation coder\\events, probes, media};
\node[artifact] (animart) at (8.8,-2.65) {Preview video\\frames + transforms};
\node[card, fill=llgreen!6] (verify) at (13.2,-1.25) {Geometry, vision,\\and video probes};
\node[card, fill=llgreen!10] (fix) at (13.2,-2.65) {Animation repair\\or LLM refinement};
\node[artifact, fill=llgreen!18, minimum width=41mm] (out) at (13.2,-4.05) {Verified Blender script and audit trail};

\draw[arrow] (prompt) -- (planner);
\draw[arrow] (planner.east) -- (scene.west);
\draw[arrow] (scene) -- (staticart);
\draw[arrow] (staticart) -- (staticfix);
\draw[arrow] (staticfix.east) -- ++(0.8,0) -- ++(0,2.8) -- (anim.west);
\draw[arrow] (anim) -- (animart);
\draw[arrow] (animart.east) -- (verify.west);
\draw[arrow] (verify) -- (fix);
\draw[arrow] (fix) -- (out);
\draw[repair] (staticfix.west) -- ++(-0.7,0) -- ++(0,2.8) -- (scene.west);
\draw[repair] (fix.west) -- ++(-0.9,0) -- ++(0,-1.0) -- (6.55,-3.65) -- (6.55,-1.25) -- (anim.west);

\begin{scope}[on background layer]
\node[draw=llblue!30, rounded corners=8pt, fill=llblue!4, fit=(s1)(prompt)(planner), inner sep=5pt] {};
\node[draw=llorange!35, rounded corners=8pt, fill=llorange!5, fit=(s2)(scene)(staticart)(staticfix), inner sep=5pt] {};
\node[draw=llpurple!35, rounded corners=8pt, fill=llpurple!5, fit=(s3)(anim)(animart), inner sep=5pt] {};
\node[draw=llgreen!35, rounded corners=8pt, fill=llgreen!5, fit=(s4)(verify)(fix)(out), inner sep=5pt] {};
\end{scope}
\end{tikzpicture}
}
\caption{The harness turns a prompt into a verified Blender animation through two verifier-repair loops. \code{SceneSpec} establishes a measured static baseline through deterministic and visual checks, and the animation extension adds temporal events, probes, media evidence, and bounded repair.}
\label{fig:architecture}
\end{figure}

\section{Related Work}

\paragraph{LLMs as program generators.}
Large language models generate code and use tools across many domains~\citep{chen2021evaluating,austin2021program,yao2023react,schick2023toolformer}. Blender scripting fits this trend because scenes, materials, cameras, render settings, and keyframes all have executable APIs. Animation raises the bar from syntactic execution to temporal correctness. Our contract makes the generated program expose the objects, events, probes, and artifacts that validation needs.

\paragraph{Code-based 3D modeling.}
VeriAnim uses LL3M code as a reference for the premise that large language models can generate Blender programs as their primary 3D output~\citep{lu2025verianim}. The system changes the problem scope from code-driven static modeling to verifier-facing animation, adding event windows, family-specific probes, video evidence, and repair contracts.

\paragraph{3D modeling IRs and scene representations.}
Existing 3D representations optimize different contracts. USD and COLLADA encode rich production assets and interchange metadata~\citep{pixar2016usd,khronos2025collada}. ShapeAssembly and StructureNet represent object part structure through programs or hierarchical graphs~\citep{jones2020shapeassembly,mo2019structurenet}. 3D scene graphs and graph-conditioned generators encode objects as nodes and relations as edges~\citep{armeni2019scenegraph,gao2024graphdreamer}. Holodeck and SceneScript show that structured scene commands can support language-guided environment generation and reconstruction~\citep{yang2024holodeck,avetisyan2024scenescript}. These IRs provide assets, layouts, or relations. \code{SceneSpec} targets a different contract: LLM-generated Blender programs must expose stable ids, collision proxies, camera evidence, verifier probes, and repair scope before animation begins.

\paragraph{Text-to-3D, text-to-motion, and video generation.}
Text-to-3D methods synthesize assets or scenes from language and image priors~\citep{poole2022dreamfusion,lin2023magic3d}. Text-to-video and text-to-4D systems generate temporal media through learned priors~\citep{ho2022imagenvideo,blattmann2023align}. These systems produce compelling pixels without editable Blender object ids, contact windows, or repairable keyframes. Our harness targets executable animation programs whose intermediate state can be inspected.

\paragraph{Motion primitives and task structure.}
Robotics systems often decompose language-conditioned action into skills, affordances, or motion primitives~\citep{huang2022inner,ahn2022can}. Classical planning and animation also emphasize constraints, paths, and motion structure~\citep{lavalle2006planning,witkin1988spacetime}. Our primitive layer borrows that decomposition, but it uses Blender geometry and media artifacts as verifier evidence. The primitive is a contract over generated keyframes, not a robot controller.

\paragraph{Physics simulation and animation plausibility.}
Physics engines provide collision response, constraints, particles, fluids, and soft bodies~\citep{coumans2016bullet,macklin2014unified}. Keyframed Blender scripts often bypass those engines because LLMs can write transforms more easily than simulation setups. We treat simulation as one possible family profile inside the extension contract. Rigid-body primitives, bounding-box probes, BVH checks, joint probes, particle statistics, and deformation probes attach to the same event structure.

\paragraph{Multimodal verification for generated content.}
Vision and video-language models judge semantic layout and temporal intent in generated media~\citep{openai2023gpt4v,li2023blip2}. Deterministic checks measure geometry and temporal predicates. The two forms of evidence answer different questions. A video model can see whether a camera orbit communicates the intended object; a deterministic probe can catch a one-frame support violation hidden by framing. The harness combines both and records their disagreement as an evaluation target.

\section{Methodology}

\subsection{Static Scene Contracts Anchor Animation}

The harness generates, verifies, repairs, and freezes a static scene before it adds animation. This staging makes the static scene a measured contract that later motion must preserve. We introduce \code{SceneSpec} as the static-scene IR for this contract. The static stage owns the object inventory, stable ids, part structure, materials, cameras, lights, environment, spatial relations, collision proxies, screenshot plan, and texture policy. The animation stage then references these ids instead of rebuilding the scene.

\code{SceneSpec} separates semantic intent from executable Blender code. \code{ObjectSpec} records controllable entities, roles, dimensions, parts, materials, and collision proxies. \code{SpatialRelationSpec} records relations such as support, attachment, containment, distance, and facing, together with a verification method. \code{CameraSpec} and \code{ScreenshotPlan} record the evidence that visual verifiers need. \code{MaterialSpec} records texture policy before code generation, so appearance retrieval remains auditable.

\begin{table}[t]
\centering
\small
\setlength{\tabcolsep}{2.5pt}
\caption{SceneSpec differs from prior 3D IRs because it is a verifier-facing contract for LLM-generated Blender programs.}
\label{tab:ir-comparison}
\resizebox{\linewidth}{!}{%
\begin{tabular}{p{0.20\linewidth}p{0.26\linewidth}p{0.23\linewidth}p{0.21\linewidth}}
\toprule
Representation & Primary contract & Missing for our setting & SceneSpec contract \\
\midrule
USD / COLLADA & Asset exchange, composition, animation channels, metadata~\citep{pixar2016usd,khronos2025collada} & No LLM planning boundary or repair evidence & Typed generation intent plus artifact ownership \\
ShapeAssembly / StructureNet & Part programs or hierarchical part graphs~\citep{jones2020shapeassembly,mo2019structurenet} & Object-level scenes, cameras, media, and verifier probes remain outside the IR & Objects, parts, roles, materials, proxies, and relations share ids \\
3D scene graphs & Semantic nodes and spatial edges~\citep{armeni2019scenegraph,gao2024graphdreamer} & Edges describe relations but do not specify how generated Blender code gets checked & Each relation names a deterministic or media verifier \\
LLM scene commands & Language-guided static environments or structured reconstruction~\citep{yang2024holodeck,avetisyan2024scenescript} & Execution traces, failure localization, and animation capability profiles are not first-class & SceneSpec freezes a repairable baseline before temporal events \\
\bottomrule
\end{tabular}
}
\end{table}

This design keeps repair local. A static support failure should move the misplaced object, not regenerate the table, camera, or material. A temporal support failure should move only the keyed path. A video verifier complaint about poor visibility should adjust camera evidence, not change object identity. The static-scene-first boundary gives every verifier and repair routine a shared baseline.

\subsection{Animation Extension Contracts Make Time Explicit}

The Animation Extension Contract is the paper's core abstraction. It extends a validated \code{SceneSpec} with four pieces of temporal structure: animation families, event windows, verifier probes, and runtime capability profiles. The same object ids appear in scene construction, generated Blender code, transform traces, deterministic reports, screenshots, preview videos, and repair prompts.

\begin{figure}[t]
\centering
\resizebox{0.96\linewidth}{!}{%
\begin{tikzpicture}[
    font=\small,
    panel/.style={draw, rounded corners=5pt, thick, fill=white, align=left, text width=36mm, inner sep=5pt},
    header/.style={draw, rounded corners=4pt, thick, align=center, minimum width=42mm, minimum height=8mm, inner sep=4pt},
    core/.style={draw, rounded corners=5pt, thick, fill=llblue!12, align=center, minimum width=48mm, minimum height=10mm},
    arrow/.style={-Latex, thick}
]
\node[core] (contract) at (0,1.75) {Animation Extension Contract};
\node[header, fill=llblue!10] (event) at (-4.4,0.35) {Event Layer};
\node[header, fill=llgreen!12] (evidence) at (0,0.35) {Evidence Layer};
\node[header, fill=llpurple!12] (artifact) at (4.4,0.35) {Artifact Layer};
\node[panel] (p1) at (-4.4,-1.05) {\textbf{Families}\\rigid, camera, visibility\\deformable, mixed};
\node[panel] (p2) at (0,-1.05) {\textbf{Verifier probes}\\bbox and BVH\\vision and video\\joint, deformation, particle stats};
\node[panel] (p4) at (4.4,-1.05) {\textbf{Artifacts and caches}\\scene graphs and traces\\frames and videos\\cache status and invalidation};
\foreach \x in {event,evidence,artifact} \draw[arrow] (contract.south) -- (\x.north);
\draw[arrow] (event) -- (p1);
\draw[arrow] (evidence) -- (p2);
\draw[arrow] (artifact) -- (p4);
\begin{scope}[on background layer]
\node[draw=llgray!25, rounded corners=8pt, fill=lllight, fit=(contract)(event)(artifact)(p1)(p4), inner sep=7pt] {};
\end{scope}
\end{tikzpicture}
}
\caption{The extension contract organizes animation into event, evidence, and artifact layers. Each layer names the information that generated Blender code must expose for verification and repair.}
\label{fig:extension-taxonomy}
\end{figure}

\textbf{Families define behavior scope.}
The current IR recognizes rigid, camera, visibility, deformable, and mixed families. Rigid motion and camera motion have executable Blender paths. Deformable behavior uses visible shape change and deformation statistics. Capability profiles record unsupported requests as scoped execution results.

\textbf{Probes define acceptance evidence.}
Each family names evidence that can validate it. Rigid events use bounding boxes, BVH checks, or convex proxies. Camera events use coverage and subject-visibility reports. Deformable events require deformation statistics and video evidence. Capability profiles report missing joint, particle, or volume probes as unsupported runtime scope.

\textbf{Capability profiles define honest execution.}
The same prompt can target Blender, Unity, Maya, or other runtimes, but the harness does not pretend that every family is executable everywhere. A target profile marks each family as supported, degraded, or unsupported. The planner and verifier report unsupported scope instead of producing artifacts that look complete but lack runtime evidence.

\subsection{Staging Preserves the Static Scene Baseline}

The harness runs a complete static verifier-repair loop before it generates animation. The static stage executes the scene script, validates object ids and spatial relations, applies deterministic support repair when measured bboxes identify a bounded placement error, renders screenshots, and runs visual verification. The animation stage references the verified ids and may add keyframes, media sampling, contact constraints, and camera events. This separation lets repair edit a motion path without rebuilding the scene to hide the original problem.

The Blender runtime acts as the shared measurement substrate. It executes generated code, records object summaries, aggregates descendants by \code{verianim\_id}, samples transforms at semantic frames and dense audit windows, renders screenshots, and renders preview videos when requested. Verifiers and repair routines consume these artifacts as structured evidence about the generated script.

\subsection{Rigid Primitives Encode Motion Contracts}

Rigid-body animation is the most mature implementation family. Its primitives are verifier-facing contracts: each helper has preconditions over roles, generated keyframes, geometric invariants, and observable predicates. The taxonomy in Figure~\ref{fig:rigid-taxonomy} organizes these contracts by the invariant they preserve, echoing earlier animation systems that expose motion structure as reusable graph or constraint elements~\citep{kovar2002motiongraphs,witkin1988spacetime}.

\begin{figure}[t]
\centering
\resizebox{0.95\linewidth}{!}{%
\begin{tikzpicture}[
    font=\small,
    level/.style={draw, rounded corners=2pt, thick, align=center, fill=white, minimum height=8mm, text width=31mm},
    root/.style={level, fill=llblue!12, text width=36mm},
    support/.style={level, fill=llgreen!12},
    interact/.style={level, fill=llorange!14},
    comp/.style={level, fill=llpurple!14, text width=46mm},
    arrow/.style={-Latex, thick}
]
\node[root] (geom) at (0,2.7) {Geometric grounding\\bbox, support top, spacing};
\node[level, fill=lllight, text width=36mm] (atomic) at (0,1.35) {Atomic kinematics\\translate, path, rotate};
\node[support] (support) at (-5.0,0) {Support motion\\slide, sequence, drop};
\node[interact] (paired) at (0,0) {Paired interaction\\push, carry};
\node[interact] (manip) at (5.0,0) {Manipulation\\pick, place, gripper};
\node[support] (artic) at (-2.5,-1.55) {Articulation\\hinge, rotor};
\node[comp] (compose) at (2.5,-1.55) {Composition rules\\one transform owner per window};
\draw[arrow] (geom) -- (atomic);
\foreach \x in {support,paired,manip} \draw[arrow] (atomic) -- (\x);
\draw[arrow] (support) -- (artic);
\draw[arrow] (paired) -- (compose);
\draw[arrow] (manip) -- (compose);
\end{tikzpicture}
}
\caption{Rigid primitives form a hierarchy of contracts. Geometry grounds object placement, atomic kinematics inserts transforms, higher-level primitives preserve support or interaction invariants, and composition rules assign one transform owner to each payload window.}
\label{fig:rigid-taxonomy}
\end{figure}

\begin{table}[t]
\centering
\caption{Rigid primitive classes and their verifier-facing contracts.}
\label{tab:primitive-taxonomy}
\begin{tabular}{p{0.22\linewidth}p{0.26\linewidth}p{0.40\linewidth}}
\toprule
Class & Helpers & Contract \\
\midrule
geometric grounding & bbox, bottom/top alignment, gripper spacing & Convert semantic roles into measured support heights, sizes, centers, and safe spacing. \\
atomic kinematics & location, rotation, scale, path keyframes & Express low-level transforms with explicit frames and interpolation. \\
support-preserving motion & support slide, support sequence, drop-to-support & Keep the subject bottom aligned to the active support top while it moves or lands. \\
paired interaction & push, attached carry & Assign a driver and payload, then maintain offset or contact over a bounded frame window. \\
manipulation lifecycle & pick-place, parallel gripper pick-place & Encode approach, grasp, lift, transfer, lower, release, and finger state from support geometry. \\
articulation and periodicity & hinge, rotor root, rotor & Rotate around a meaningful pivot while preserving the mechanism body. \\
composition rules & event windows, role tags, static checks & Prevent overlapping transform ownership and preserve the verified scene baseline. \\
\bottomrule
\end{tabular}
\end{table}

The taxonomy is motivated by the structure of Blender animation, not by a catalog of bugs. Language prompts usually express intent through relations such as on, inside, holding, attached, following, and opening. Each relation maps to a different invariant. Support motion owns vertical placement; carry motion owns a driver-payload offset; gripper manipulation owns a phase sequence; hinged motion owns an origin and axis; composition owns time windows. Development traces helped calibrate tolerances and static checks, but the taxonomy follows these invariants.

\subsection{Verification Combines Geometry and Media}

The deterministic verifier checks structured predicates over sampled frames. Support predicates require footprint overlap and a small vertical gap. Nonpenetration predicates measure overlap depth across collision-enabled pairs. Carry-contact predicates allow intended proximity while detecting embedding. Containment predicates check whether centers or bounding boxes remain inside a container proxy. Dense audit windows catch interpolation violations that sparse start/middle/end samples miss.

Media verifiers answer questions that geometry cannot answer alone. Vision checks judge static layout, material intent, relation visibility, and camera framing. Video checks judge temporal order, apparent motion, final state, and subject visibility. The harness sends video artifacts, ordered sampled frames, transform traces, and deterministic reports together, so multimodal feedback can point at the same object ids and frame windows as deterministic evidence.

\subsection{Repair Uses Measured Evidence}

Repair routines edit local scene placement or motion when the verifier reports bounded evidence. Static support repair recalculates object placement from the support top before animation begins. Temporal support repair recalculates the subject's vertical coordinate at keyed frames. Path repair shifts keyed locations in the failing window. Bridge and platform repair create outside-lift-enter-cross-exit paths. Carry repair sends role-specific evidence to the refiner when a generated script violates gripper or payload ownership. The repair loop stays bounded: repeated identical failure signatures stop refinement instead of running indefinitely.

\section{System Components}

The implemented harness includes a Blender addon server, Python runner, typed IR parser, texture resolver, visual verifier, video verifier, deterministic validator, and refinement loop. The addon exposes commands for code execution, scene inspection, material inspection, camera coverage reports, screenshot rendering, animation sampling, and preview video rendering. Each run stores IR snapshots, generated scripts, refined scripts, verifier reports, media paths, selected textures, transform traces, and model configuration.

Texture resolution is an explicit preprocessing step. The planner marks materials with \code{texture\_policy}, \code{needs\_texture}, and \code{texture\_query}. A material agent searches public texture pages, downloads candidates, asks a vision model to approve surface suitability, and writes approved local paths back into the IR. If search or vision approval fails, the coder falls back to procedural colors and shader parameters. This keeps appearance assets auditable without making texture retrieval part of the animation claim.

Prompt caching supports iterative refinement. The stable \code{verianim\_utils} API contract appears in a fixed prompt prefix shared by coder, refiner, user-edit, and animation-extension calls. Per-run IR, verifier reports, and media evidence remain in the user payload. This organization reduces repeated prompt cost while preserving fresh failure evidence.

\section{Benchmark and Evaluation}

VeriAnim-AnimBench tests both the contract and the current primitive library. Figure~\ref{fig:evaluation-plan} organizes the evaluation matrix. Prompt families span rigid support motion, paired interaction, manipulation, articulation, camera motion, visibility/state changes, deformable events, and mixed scenes. System variants remove one layer at a time: direct keyframes, IR without extension profiles, extension without primitives, primitives without repair, and the full verifier-facing harness.

We define \emph{VeriAnim-AnimBench}, a 300-prompt benchmark stored in \code{benchmark/}. The benchmark contains 100 easy, 100 medium, and 100 hard natural-language prompts. Easy prompts isolate one object or one relation, medium prompts combine two coordinated events or camera requirements, and hard prompts require multi-object composition, concurrent motion, or family mixtures. Each record stores the prompt, animation families, required motions, verifier focus, and expected difficulty rationale.

\begin{figure}[t]
\centering
\small
\setlength{\tabcolsep}{2.5pt}
\resizebox{0.96\linewidth}{!}{%
\begin{tabular}{p{0.19\linewidth}p{0.13\linewidth}p{0.12\linewidth}p{0.14\linewidth}p{0.15\linewidth}p{0.14\linewidth}}
\toprule
Family / variant & Direct keys & IR only & Contract & + primitives & + repair \\
\midrule
support + carry & weak & ids & probes & contacts & local fix \\
pick-place & weak & events & probes & lifecycle & ownership \\
hinge / rotor & axis & events & pivot & helpers & checks \\
camera orbit & path & target & coverage & helper & media \\
deformable & shape & event & stats & deformation & media \\
\bottomrule
\end{tabular}
}
\caption{The evaluation matrix tests animation families against staged system variants. The main comparison asks whether contracts, primitives, and repair each improve measurable validity and media-verified temporal intent.}
\label{fig:evaluation-plan}
\end{figure}

\textbf{Metrics.}
The primary metrics are deterministic pass rate, support violation rate, penetration rate, final relation success, transform-ownership violations, subject visibility, video verifier pass rate, number of refinement rounds, and wall-clock execution cost. Deformable events add deformation-statistic pass rate. Capability-profile cases report missing joint, particle, or volume probes as unsupported scope.

\textbf{Ablations.}
The evaluation isolates the contribution of each layer. Removing the extension contract tests whether a flat animation spec provides enough structure. Removing primitives tests whether LLM-authored keyframes satisfy the same event contracts. Removing dense-frame audits tests whether sparse samples miss interpolation errors. Removing video verification tests whether deterministic geometry alone captures temporal intent and camera coverage.

\textbf{Independent audits.}
Bounding boxes share assumptions with several rigid primitives, so the benchmark includes independent audit hooks. Human ratings judge visual plausibility and semantic success. Mesh-level BVH checks audit rigid scenes. Deformation statistics validate shape change beyond video impressions. Disagreement among deterministic, media, and human judgments becomes a reported result for verifier design.

\section{Limitations}

The current system provides a contract layer and a runnable Blender harness, while full physics simulation remains outside its scope. Rigid primitives are kinematic and do not model forces, friction, mass, torque, or collision response. Bounding boxes over-approximate concave objects and under-specify fine mesh intersections. Deformable events use statistics and video evidence in place of a full cloth solver. Character, mocap, IK, fluid, smoke, and particle requests are reported through capability profiles when runtime probes are unavailable. Video verification helps judge temporal intent, but deterministic probes remain necessary for short contact violations and hidden geometry errors.

\section{Conclusion}

Animation changes VeriAnim-style Blender generation from static modeling into temporal program synthesis. The Animation Extension Contract makes that shift explicit: it freezes a verified scene baseline, adds typed animation families, names verifier probes, records runtime capability, and routes repair through measured evidence. Rigid-body primitives demonstrate the contract's value today by turning support, carry, pick-place, articulation, and composition into geometry-aware Blender helpers. The broader methodology replaces direct keyframe writing with staged, auditable, and extensible animation contracts.

\bibliography{references}
\bibliographystyle{iclr2026_conference}

\end{document}
